### 2014-01-30 inst/doc/datasets.Rmd

# ReAnalysis based on reviewer comments

Because we have re-worked the publication as a paper about the tool, we have decided to do a re-analysis of the original data presented in the original manuscript based on the reviewer comments, as well as include other data sets.

# Muscle VS Skin Denervation

The original data consisted of muscle denervation data from GSE4411, and skin denervation data from Dr. Jeffrey Petruska.

# UC VS Crohn's

From GEO there is a ulcerative colitis vs crohn's disease data set available, GSE36807. This would be a nice one to try, as there is already a publication on it.

# Metabolomics VS Transcriptomics

The NCI 60 database has both metabolomics and transcriptomics data available. Unfortunately, there is no **control** data set.

Metabolomics File:

  * File Format: SAMPLENAME, TITLE, PANELNBR, CELLNBR, pname, cellname, VALUE

Ideas for analysis: slow vs fast growing, sensitive to compound vs insensitive to compound

Transcriptomics File:

  * File Format: MOLTID (NCI pattern #), ACC, Gene, TITLE, MOLTNBR (NCI exp. id #), PANELNBR, CELLNBR, pname, cellname, CHIP, FEATURE_ID, UniGene, GeneID, VALUE Gene assignments are based on Unigene Build #U225 (August 2010)