### 2014-01-30 inst/doc/process_lung_data.Rmd

<!--
  %\VignetteEngine{knitr}
%\VignetteIndexEntry{process lung data}
%\VignetteDepends{GEOquery}
-->

```{r customCSS, include=FALSE}
cssFile <- system.file("extdata", "style.css", package="ccPaperRev")
options(markdown.HTML.stylesheet = cssFile)
```

# Lung Data

One of the experimental data sets we use is from [GEO][http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE18842]. This documents the actual processing.

```{r downloadData, eval=FALSE}
download.file("ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE18nnn/GSE18842/matrix/GSE18842_series_matrix.txt.gz", "/mlab/data/rmflight/Documents/projects/work/categoryComparePaperRev/otherData/GSE18842_series_matrix.txt.gz")
```

```{r loadResave, eval=FALSE}
library(GEOquery)
lung <- getGEO(filename="otherData/GSE18842_series_matrix.txt.gz")
tmpP <- pData(lung)
tmpP <- tmpP[,c("title", "geo_accession", "platform_id", "data_processing", "supplementary_file", "source_name_ch1", "characteristics_ch1", "characteristics_ch1.1")]
pData(lung) <- tmpP

tmpF <- fData(lung)
tmpF <- tmpF[, c("ID", "GB_ACC", "SPOT_ID", "Annotation Date", "Sequence Type", "Sequence Source", "Target Description", "Representative Public ID", "Gene Title", "Gene Symbol", "ENTREZ_GENE_ID", "RefSeq Transcript ID")]
fData(lung) <- tmpF
save(lung, file="inst/data/lung.RData", compress="xz")
```