### 2014-01-30 inst/doc/uc_crohns_analysis.Rmd

<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{UC vs Crohn's}
%\VignetteDepends{ccPaperRev, GEOquery, limma, hgu133plus2.db, GO.db, org.Hs.eg.db}
-->

```{r customCSS, include=FALSE}
cssFile <- system.file("extdata", "style.css", package="ccPaperRev")
options(markdown.HTML.stylesheet = cssFile)
```

```{r loadLibrary, message=FALSE}
library(ccPaperRev)
```

# Crohn's VS Ulcerative Colitis

GEO dataset **GSE36807** has crohn's, uc, and normal samples.

```{r getData, eval=FALSE}
library(GEOquery)

cucData <- getGEO(GEO="GSE36807")[[1]]

tmpP <- pData(cucData)
tmpP$status <- "control"

cLoc <- grep("Crohn", tmpP$characteristics_ch1)
tmpP$status[cLoc] <- "crohns"

ucLoc <- grep("Ulcerative colitis", tmpP$characteristics_ch1)
tmpP$status[ucLoc] <- "uc"
pData(cucData) <- tmpP

.sessionInfo <- sessionInfo()
.timeDate <- Sys.time()
save(cucData, .sessionInfo, .timeDate, file="inst/data/uc_crohns_rawData.RData")
```

## Differential analysis

Use the `limma` package to rank probes.

```{r rankProbes}
library(limma)
library(hgu133plus2.db)
data(uc_crohns_rawData)

cucComps <- c("uc - control", "crohns - control")
geneID <- unlist(mget(featureNames(cucData), hgu133plus2ENTREZID))

cucExpr <- exprs(cucData)
cucCollapse <- collapseProbes(cucExpr, geneID) # collapse to single genes using median of expression
cucCharacter <- pData(cucData)

cucFC <- rankGenes(cucCollapse, cucCharacter$status, cucComps, doAggregation=FALSE, aggregateIndex=NA)
names(cucFC) <- c("UC", "CROHNS")

cucDiff <- lapply(cucFC, getDiffGenes, id="id")
# this object can be run using base ccEnrich, or can be easily coerced
```

## ORA

Do normal over-representation analysis on the `cucDiff`, those genes differentially expressed in the Crohn's and UC samples.

```{r ORA}
library(GO.db)
cucGeneList <- list(UC_up=list(genes=cucDiff$UC$up, universe=cucDiff$UC$universe, annotation="org.Hs.eg.db"),
                    UC_dn=list(genes=cucDiff$UC$dn, universe=cucDiff$UC$universe, annotation="org.Hs.eg.db"),
                    CR_up=list(genes=cucDiff$CROHNS$up, universe=cucDiff$CROHNS$universe, annotation="org.Hs.eg.db"),
                    CR_dn=list(genes=cucDiff$CROHNS$dn, universe=cucDiff$CROHNS$universe, annotation="org.Hs.eg.db"))
cucGeneList <- new("ccGeneList", cucGeneList, ccType="BP")

cucEnrich <- ccEnrich(cucGeneList)
pvalueType(cucEnrich) <- "pval"
pvalueCutoff(cucEnrich) <- 0.01

cucEnrich

cucOpts <- new("ccOptions", listNames=names(cucGeneList), compareNames=
                c("UC_up", "UC_dn", "CR_up", "CR_dn", "UC_up,CR_up", "UC_up,CR_dn", "UC_dn,CR_up", "UC_dn,CR_dn"))
cucCompare <- ccCompare(cucEnrich, cucOpts)

cucCompare

```

```{r out2Cytoscape, eval=FALSE}
cucCompare <- breakEdges(cucCompare$BP, 0.8)
cucCy <- ccOutCyt(cucCompare, cucOpts)
breakEdges(cucCy, 1)
```

## Run GSEA type analysis using limma

There are two ways to do GSEA using `limma`.

  1 Use `romer` for competitive tests with rotations to do random samples
  2 Use `geneSetTest` for gene-wise competitive tests, *i.e.* each set is tested against random gene samples

We're going to try both and see what we get out.

```{r setupGeneSets}
library(GO.db)
library(org.Hs.eg.db)
library(limma)
hsGO <- as.list(org.Hs.egGO2ALLEGS)
hsGO <- hsGO[(Ontology(names(hsGO)) == "BP")]
hsGO <- lapply(hsGO, unique)

useSet <- symbols2indices(hsGO, rownames(cucCollapse))
```

Run using `romer`

```{r setupContrasts}
sampleStatus <- cucCharacter$status
doComps <- cucComps
f <- factor(sampleStatus)
design <- model.matrix(~0 + f)
colnames(design) <- levels(f)

contrast.matrix <- makeContrasts(contrasts=doComps, levels=design)
```

```{r runRomer, eval=FALSE}
cucRomer <- multicontrastRomer(useSet, cucCollapse, design, contrast.matrix, nrot=10000)
names(cucRomer) <- c("UC", "CROHNS")
save(cucRomer, file="inst/data/cucRomer.RData")

```

```{r examineResults}
data(cucRomer)

geneAnnMapping <- new("namedList", .Data=hsGO, names=names(hsGO))

getSigID <- function(inRomer, pCut=0.05, whichCol=c("Up", "Down")){
  sigID <- lapply(whichCol, function(inCol){
    rownames(inRomer[(inRomer[, inCol] <= pCut),])
  })
  names(sigID) <- whichCol
  return(sigID)
}

cucSigRomer <- lapply(cucRomer, getSigID, pCut=0.01)
cucSigRomer <- unlist(cucSigRomer, recursive=FALSE)

genCCSigList <- function(inSig){
  tmp <- new("ccSigList", sigID=inSig)
}

cucCCSig <- lapply(cucSigRomer, genCCSigList)

cucCCRomer <- new("GENccEnrichResult", cucCCSig, categoryName="GSEAGO", geneAnnMapping=geneAnnMapping, overlapType="overlap", annDescription=Term(names(geneAnnMapping)))

cucRomerOpts <- new("ccOptions", listNames=names(cucCCRomer), compareNames=c(
  "UC.Up", "UC.Down", "CROHNS.Up", "CROHNS.Down", "UC.Up,CROHNS.Up", "UC.Up,CROHNS.Down", "UC.Down,CROHNS.Up", "UC.Down,CROHNS.Down"))
compareCUCRomer <- ccCompare(cucCCRomer, cucRomerOpts)
compareCUCRomer
```

```{r send2Cytoscape, eval=FALSE}
compareCUCRomer <- breakEdges(compareCUCRomer, 0.8)
cucRomerCW <- ccOutCyt(compareCUCRomer, cucRomerOpts, postText="romer", rpcPort=9001)
generateLegend(cucRomerOpts)
```